An efficient acyclic contact planner for multiped robots

### 2016-01-27 body/results.tex
nicolas comments

In our previous work~\citep{tonneauisrr15} additional results are demonstrated with various virtual avatars (Figure~\ref{fig:robots_old}).
In this extension we choose to focus on actual robots. We invite the interested reader to watch the ISRR video\endnote{https://www.youtube.com/watch?v=LmLAHgGQJGA}, and 
to refer to the previous paper for a discussion on these results.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/stair}
  \caption{
           HRP-2 in the steep stair climbing scenario. }
		   \label{fig:stair_robust}
\end{figure*}

@@

\noindent\textbf{Heuristics:} The manipulability $h_w$ is chosen for the feet; $h_{EFORT}$ is chosen for the right arm.
Regarding equilibrium, the video demonstrates two sequences computed for two different threshold values of $b_0$: $0$ and $2$ (Figure~\ref{fig:stair_robust}).

@@

Again, the interpolation motion between the contact steps is out of the scope of this paper. However it should be noted that the computed plan in this scenario has been executed successfully on the robot~\citep{Carpentier2016}.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/standing}
  \caption{
           HRP-2 in the standing scenario. }
		   \label{fig:standing}
\end{figure*}

@@

\noindent\textbf{Heuristics:} $h_w$ for the feet, $h_{EFORT}$  for the hands.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/darpa}
  \caption{
           Robust crossing of rubbles by HyQ ($b_0 > 20$). }
		   \label{fig:darpa}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/HyQ_obs}
  \caption{
           Crossing a hole contact sequence for HyQ ($b_0 > 4$). }
		   \label{fig:HyQ_obs}
\end{figure*}

@@

\noindent\textbf{Heuristics:} $h_w$  for all legs. $h_{EFORT}$  for the right arm. The robustness threshold is set to $2$.

@@

\noindent\textbf{Heuristics:} $h_{EFORT}$ for all fingers.

@@

\subsubsection{Choosing the scaling factor $s$} \label{sec:params}
To find a convenient value $s^*$ of $s$, we proceeded as follows. 
For several values of $s$, we generated $10 000$ configurations. 
We then computed the rates of false positives (i.e. the configuration satisfies
the reachability condition, but does not belong to \gls{ccequil0}) and false negatives (i.e. the configuration does not satisfy the 
reachability condition, but belongs to \gls{ccequil0}).

@@

\subsubsection{On the choice of heuristics} \label{sec:heuristichoices}
In our conference paper, the computed motions were only generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
Regarding locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation on the real robot turned out to be harder \citep{Carpentier2016}.
For this reason, we now prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, that results in less contact repositioning.

\subsubsection{On the robustness equilibrium criterion} \label{sec:parrob}
Robustness is really important when considering practical applications on the robot, to account 
for the various uncertainties that result from environment and state estimation. However,
maximizing the robustness criterion is not necessarily optimal, because the resulting configurations may be too conservative, thus not favoring the motion. In our experiments, we choose not to maximize the robustness,
but to empirically set a robustness threshold value under which a configuration is not considered to be in
static equilibrium (If no candidate reaches the threshold, instead of failing, the algorithm can eventually return the ``more robust'' configuration
found). Currently for HRP-2, a threshold value of 2 subjectively gives the best results in the considered scenarios, while 10 seems to be a good choice for HyQ. Both values do not
significantly slow down the planning times.

\subsubsection{Discretization of the guide path} \label{sec:disc}
Currently, to discretize the path, the user defines a fixed step size. The step size
has an influence on the output of the planner: if too large steps are taken,
the planner may fail since we impose the constraint that only one contact change might occur
between two consecutive steps. For instance, in most scenarios the torso of HRP-2 moves about 15 cm between two postures, but only 3 cm
for the car egress scenario, where we had to lower the step size because of the many contact breaks induced by the extremely constraining geometry.
For future work, we would like to automatically adapt the size of the discretization step to the complexity of the planning, for instance
by reducing the step size when failure occurs.

\subsection{Performance analysis} \label{sec:perf}
To analyze performance, for each considered scenario, we ran the simulation 1000 times.
We measured the computation time spent in each aspect of the algorithm, and also analyzed the success
rate obtained for each scenario.

@@

\subsubsection{Computation times}
Table~\ref{tab:requestime} summarizes the performance measurements obtained, in terms of computation times.

Considering the repartition of the computation time, for HRP-2, most of the time is spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the car egress scenario,
the kinematic constraints are really strong. Therefore a large number of projections fail because of joint-limit violations.

@@

\subsubsection{Success rates}
Table~\ref{tab:requestpercent} summarizes the success rates obtained for each scenario.
We observe that, as expected, our planner does not succeed systematically, because of the approximations made in our formulation. The extreme situation of the car egress scenario provides the less satisfying results in terms of kinematic failure. This is not
surprising considering the narrowness of the scene.
From a pragmatic point of view, regarding the computation times, we claim that our approach provides a satisfying compromise between completeness and efficiency.
Indeed, the advantage of the framework is that when the contact generation fails, it does so rapidly, which allows us to rapidly re-plan a new contact sequence with a reasonable chance of success.
The most efficient (and immediate) approach is probably to launch in parallel several instances of the planner for a given problem (our current implementation is single threaded) and to apply any successful result.

\begin{table*}
\centering
\begin{tabular}{ c | c | c | c | c }
 Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{5.42 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours} + \citeauthor{Carpentier2016}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & $ <$ 2s \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{4.08 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & X \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{10.08 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
 \hline
   \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{3.5 hours}  \\
							 & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 21 s} & X \\
 \end{tabular}
\caption{TODO VERIFIE STATS A LA MAISON.}
\label{tab:compprev}
\quad
\end{table*}
\subsection{Comparison with previous work} \label{sec:compa}
Comparing our method with others is not trivial. Indeed, we focus on the contact planning aspect of the problem, and do not address
the interpolation in this paper, nor provide a strong guarantee that our sequences can be interpolated.
Most papers in the literature only provide the overall computation
time, without specifying how much is spent in each phase. Due to the complexity
of the existing approaches,  we cannot afford the development resources required to reimplement them.

>>>>

In our previous work~\citep{tonneauisrr15} additional results are demonstrated with various virtual avatars (Figure~\ref{fig:robots_old}).
In this extension we choose to focus on actual robots. We invite the interested reader to watch the ISRR video (\url{http://youtu.be/LmLAHgGQJGA}), and 
to refer to the previous paper for a discussion on these results.

@@

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/stair}
  \caption{
           HRP-2 in the steep stair climbing scenario. }
		   \label{fig:stair_robust}
\end{figure}

@@

\noindent\textbf{Heuristics:} The manipulability $h_w$ is chosen for the feet; $h_{\textrm{\it EFORT}}$ is chosen for the right arm.
Regarding equilibrium, the video demonstrates two sequences computed for two different threshold values of $b_0$: $0$ and $2$ (Figure~\ref{fig:stair_robust}).

@@

Again, the interpolation motion between the contact steps is out of the scope of this paper. However it should be noted that the computed plan in this scenario has been executed successfully on the robot (\url{http://youtu.be/YjL-DBQgXwk#t=0m28s}).

@@

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/standing}
  \caption{
           HRP-2 in the standing scenario. }
		   \label{fig:standing}
\end{figure}

@@

\noindent\textbf{Heuristics:} $h_w$ for the feet, $h_{\textrm{\it EFORT}}$  for the hands.

@@

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/darpa}
  \caption{
           Robust crossing of rubbles by HyQ ($b_0 > 20$). }
		   \label{fig:darpa}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/HyQ_obs}
  \caption{
           Crossing a hole contact sequence for HyQ ($b_0 > 4$). }
		   \label{fig:HyQ_obs}
\end{figure}

@@

\noindent\textbf{Heuristics:} $h_w$  for all legs. $h_{\textrm{\it EFORT}}$  for the right arm. The robustness threshold is set to $2$.

@@

\noindent\textbf{Heuristics:} $h_{\textrm{\it EFORT}}$ for all fingers.

@@

\subsubsection{Choosing the scaling factor $s$:} \label{sec:params}
To find a convenient value $s^*$ of $s$, we proceeded as follows. 
For several values of $s$, we generated $10 000$ configurations. 
We then computed the rates of false positives (i.e. the configuration satisfies
the reachability condition, but does not belong to \gls{ccequil0}) and false negatives (i.e. the configuration does not satisfy the 
reachability condition, but belongs to \gls{ccequil0}).

@@

\subsubsection{On the choice of heuristics:} \label{sec:heuristichoices}
In our conference paper, the computed motions were only generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
Regarding locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation on the real robot turned out to be harder \citep{Carpentier2016}.
For this reason, we prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, that results in less contact repositioning.

\subsubsection{On the robustness equilibrium criterion:} \label{sec:parrob}
Robustness is really important when considering practical applications on the robot, to account 
for the various uncertainties that result from environment and state estimation. However,
maximizing the robustness criterion is often not optimal, because the resulting configurations may be too conservative, thus not favoring the motion. In our experiments, we choose not to maximize the robustness,
but to empirically set a robustness threshold value under which a configuration is not considered to be in
static equilibrium (If no candidate reaches the threshold, instead of failing, the algorithm can eventually return the ``more robust'' configuration
found). Currently for HRP-2, a threshold value of 2 subjectively gives the best results in the considered scenarios, while 10 seems to be a good choice for HyQ. Both values do not
significantly slow down the planning times.

\subsubsection{Discretization of the guide path:} \label{sec:disc}
Currently, to discretize the path, the user defines a fixed step size. The step size
has an influence on the output of the planner: if too large steps are taken,
the planner may fail since we impose the constraint that only one contact change might occur
between two consecutive steps. For instance, in most scenarios the torso of HRP-2 moves about 15 cm between two postures, but only 3 cm
for the car egress scenario, where we had to lower the step size because of the many contact breaks induced by the extremely constraining geometry.
For future work, we would like to automatically adapt the size of the discretization step to the complexity of the planning, for instance
by reducing the step size when failure occurs.

\subsection{Performance analysis} \label{sec:perf}
To analyze performance, for each considered scenario, we ran the planner 1000 times.
We measured the computation time spent in each aspect of the algorithm, and also analyzed the success
rate obtained for each scenario.

@@

\subsubsection{Computation times:}
Table~\ref{tab:requestime} summarizes the performance measurements obtained, in terms of computation times.

Considering the repartition of the computation time, for HRP-2, most of the time is spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the car egress scenario,
the kinematic constraints are very demanding. Therefore a large number of projections fail because of joint-limit violations.

@@

\subsubsection{Success rates:}
Table~\ref{tab:requestpercent} summarizes the success rates obtained for each scenario.
We observe that, as expected, our planner does not succeed systematically, because of the approximations made in our formulation. The extreme situation of the car egress scenario provides the less satisfying results in terms of kinematic failure. This is not
surprising considering the narrowness of the scene.
From a pragmatic point of view, regarding the computation times, we claim that our approach provides a satisfying compromise between completeness and efficiency.
Indeed, the advantage of the framework is that when the contact generation fails, it does so rapidly, which allows us to rapidly re-plan a new contact sequence with a reasonable chance of success.
The most efficient (and immediate) approach is probably to launch in parallel several instances of the planner for a given problem (our current implementation is single threaded) and to apply any successful result.

\begin{table*}
\centering
\begin{tabular}{ c | c | c | c | c }
 Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}5.42 min}  \\
							 & \textbf{Ours} + \citeauthor{Carpentier2016}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & $ <$ 2s \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}4.08 min}  \\
							 & \textbf{Ours}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}10.08 min}  \\
							 & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{\cellcolor{blue!25}3.5 hours}  \\
							 & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 21 s} & X \\
 \end{tabular}
\caption{Comparison between the computation times obtained by our method and previous ones, for each phase of the planning. Blue cells indicate merged times (because no
precise time information is available). The X symbol indicates scenarios where we did not address the interpolation of the contact sequence.}
\label{tab:compprev}
\quad
\end{table*}
\subsection{Comparison with previous work} \label{sec:compa}
Comparing our method with others is not trivial. Indeed, we focus on the contact planning aspect of the problem, and do not address
the interpolation in this paper, nor provide a strong guarantee that our sequences can be interpolated.
Most papers in the literature only provide the overall computation
time, without specifying how much is spent in each phase. Moreover, there exists no off-the-shelf implementation
of the related methods. Due to their complexity, we cannot afford the development resources required to reimplement them.
